Stellar Radiation is Critical for Regulating Star Formation and Driving Outflows in Low Mass Dwarf Galaxies

### 2019-03-18 thesis/chapters/ch_chapter3.tex
Finalized draft updates to nucleosynthesis section. Adding new references. Removing redundant references. Renaming Ji+2018 references

\subsection{Implications for Exotic Enrichment Sources}
\label{sec:exotic enrichment}
%   AE: Benoit suggested removing arguments relating to NS-NS merger energetics as I had misred / miscaclulated the energy release values. He said greater emphasise on potential differences of mixing due to rarity, rather than energy.
We have shown that the dynamical evolution of metals depends upon their nucleosynthetic source, focusing here on AGB synthesized elements and SN synthesized elements. However, these differences should also apply to exotic enrichment sources, such as hypernovae, neutron-star neutron-star mergers, and neutron-star black-hole mergers. These sources can have energies that differ significantly from typical SNe, reaching $> 10^{52}$~erg for hypernovae \citep{Nomoto2004}, for example. In addition, 
%mm the rarity of these events, for example,  there are approximately $10^{-5}$ per M$_{\odot}$ of star formation NSM in the Milky Way, could
     these events are rare.  For example, neutron star mergers occur 
     at a rate of approximately $10^{-5}$ per M$_{\odot}^{-1}$ of star
     formation in the Milky Way\citep{Kim2015}.  This rarity could 
significantly influence the typical ISM environments into which they eject their metals. We would therefore expect different mixing behaviors for metals synthesized through these channels as compared to SNe. Based on our results here, for example, we would expect elements from hypernovae to be more well mixed in the ISM of all galaxies, but more readily ejected in dwarf galaxies, as compared to elements from SNe. 
% Benoit suggesed removing this given energy argument removal above
%    Elements from NSMs, a significant source of r-process enrichment, would be less well-mixed in the ISM of all galaxies and more easily retained than SNe elements in dwarf galaxies. 
Depending on their injection energy, neutron star mergers could exhibit different mixing behaviors in the ISM. These differences could provide important signatures for distinguishing individual, exotic enrichment sources from observed stellar abundance patterns in our own Milky Way and in nearby dwarf galaxies. In particular, low mass dwarf galaxies with unusual (compared to Milky Way, solar abundances) r-process enrichment \citep[e.g.][]{Ji2016,Ji2018,Duggan2018}, are valuable for constraining the source of these elements, their frequency, and typical yields. The differential metal evolution presented in our work both opens up an additional avenue by which elements from distinct nucleosynthetic sources may be distinguished in observations and challenges current assumptions used in interpreting these observations.

>>>>

\subsection{Implications for Exotic Enrichment Sources}
\label{sec:exotic enrichment}
%   AE: Benoit suggested removing arguments relating to NS-NS merger energetics as I had misred / miscaclulated the energy release values. He said greater emphasise on potential differences of mixing due to rarity, rather than energy.
We have shown that the dynamical evolution of metals depends upon their nucleosynthetic source, focusing here on AGB synthesized elements and SN synthesized elements. However, these differences should also apply to exotic enrichment sources, such as hypernovae, neutron-star neutron-star mergers, and neutron-star black-hole mergers. These sources can have energies that differ significantly from typical SNe, reaching $> 10^{52}$~erg for hypernovae \citep{Nomoto2004}, for example. In addition, 
%mm the rarity of these events, for example,  there are approximately $10^{-5}$ per M$_{\odot}$ of star formation NSM in the Milky Way, could
     these events are rare.  For example, neutron star mergers occur 
     at a rate of approximately $10^{-5}$ per M$_{\odot}^{-1}$ of star
     formation in the Milky Way\citep{Kim2015}.  This rarity could 
significantly influence the typical ISM environments into which they eject their metals. We would therefore expect different mixing behaviors for metals synthesized through these channels as compared to SNe. Based on our results here, for example, we would expect elements from hypernovae to be more well mixed in the ISM of all galaxies, but more readily ejected in dwarf galaxies, as compared to elements from SNe. 
% Benoit suggesed removing this given energy argument removal above
%    Elements from NSMs, a significant source of r-process enrichment, would be less well-mixed in the ISM of all galaxies and more easily retained than SNe elements in dwarf galaxies. 
Depending on their injection energy, neutron star mergers could exhibit different mixing behaviors in the ISM. These differences could provide important signatures for distinguishing individual, exotic enrichment sources from observed stellar abundance patterns in our own Milky Way and in nearby dwarf galaxies. In particular, low mass dwarf galaxies with unusual (compared to Milky Way, solar abundances) r-process enrichment \citep[e.g.][]{Ji2016,Ji2018b,Duggan2018}, are valuable for constraining the source of these elements, their frequency, and typical yields. The differential metal evolution presented in our work both opens up an additional avenue by which elements from distinct nucleosynthetic sources may be distinguished in observations and challenges current assumptions used in interpreting these observations.

### 2019-03-18 thesis/chapters/ch_introduction.tex
Finalized draft updates to nucleosynthesis section. Adding new references. Removing redundant references. Renaming Ji+2018 references

This means that elements from different nucleosynthetic channels are produced on different timescales during a galaxy's evolution, as set by the lifetimes of the stars in which they occur, and are released into the galaxy at different energies, with different spatial distributions within the galaxy, and at different rates. For example, O, an $\Alpha$-element is produced predominately in massive stars and is therefore released via core collapse supernovae on timescales of $\sim$~10~Myr after a star formation event, near their stellar birth sites. By contrast, Sr is produced predominately via the s-process (discussed more below) which occurs in the asymptotic giant branch (AGB) winds of lower mass stars ($M < 8$~M$_{\odot}$) at much lower energies and on timescales of $\sim 100$~Myr - $1$~Gyr. By this time, these stars would have diffused farther away from their birth sites, potentially sampling a qualitatively different portion of the ISM than earlier core collapse SNe. Generally the differences in how different timescales affect observed stellar abundances patterns are well understood and easily modelled.\footnote{The canonical example of this is the ``knee" feature in a $[\Alpha/$Fe] vs. [Fe/H] diagram, where the location of this knee signifies the point at which Type Ia enrichment (which produces significant Fe, but little to no $\Alpha$ elements) begins to have a significant contribution over core collapse SNe (rich in $\Alpha$ elements, but not Fe) \citep[e.g.[]{MatteucciBrocato1990, Geisler2005, Hill2018}. The location of this knee shifts in a galaxy depending on its SFH.} However, it is not well understood how differences in where and with what energies these events occur affect observed stellar abundance patterns. Investigating these differences is a key motivation of this dissertation.

As relevant to this work, we briefly summarize a few of the important nucleosynthetic channels shown in Figure~\ref{intro:fig:periodic table} and their sites. $\Alpha$-elements, notably O, Mg, Si, Ca, and Ti, are produced during the lifetime of massive stars (O, Mg) or during the core collapse supernova explosion itself. In either case, these elements are

>>>>

This means that elements from different nucleosynthetic channels are produced on different timescales during a galaxy's evolution, as set by the lifetimes of the stars in which they occur, and are released into the galaxy at different energies, with different spatial distributions within the galaxy, and at different rates. For example, O, an $\alpha$-element is produced predominately in massive stars and is therefore released via core collapse supernovae on timescales of $\sim$~10~Myr after a star formation event, near their stellar birth sites. By contrast, Sr is produced predominately via the s-process (discussed more below) which occurs in the asymptotic giant branch (AGB) winds of lower mass stars ($M < 8$~M$_{\odot}$) at much lower energies and on timescales of $\sim 100$~Myr - $1$~Gyr. By this time, these stars would have diffused farther away from their birth sites, potentially sampling a qualitatively different portion of the ISM than earlier core collapse SNe. Generally the differences in how different timescales affect observed stellar abundances patterns are well understood and easily modelled.\footnote{The canonical example of this is the ``knee" feature in a $[\alpha/$Fe] vs. [Fe/H] diagram, where the location of this knee signifies the point at which Type Ia enrichment (which produces significant Fe, but little to no $\alpha$ elements) begins to have a significant contribution over core collapse SNe (rich in $\alpha$ elements, but not Fe) \citep[e.g.[]{MatteucciBrocato1990, Geisler2005, Hill2018}. The location of this knee shifts in a galaxy depending on its SFH.} However, it is not well understood how differences in where and with what energies these events occur affect observed stellar abundance patterns. Investigating these differences is a key motivation of this dissertation.

As relevant to this work, we briefly summarize a few of the important nucleosynthetic channels shown in Figure~\ref{intro:fig:periodic table} and their sites. $\alpha$-elements, notably O, Mg, Si, Ca, and Ti, are produced during the lifetime of massive stars (O, Mg) or during the core collapse supernova explosion itself. In either case, these elements are synthesized through the capture of an $\alpha$ nucleus, progressing towards increasingly heavier nuclei in intervals of 4/2 in mass/atomic number (as $\alpha$ = $^{4}_2He$) from $^8_4Be$ to the stable $^{52}_{26}Fe$ and unstable $^{56}_{28}Ni$ (which days to $^{56}_{26}Fe$). The iron group elements, notably Fe, Mn, Co, and Ni, are produced predominately in Type Ia supernovae, with variations depending on a single vs. double degenerate scenario and the initial composition of exploding white dwarf(s) (WD).

The presence of these heavier elements in stars enables the formation of even heavier nuclei through neutron capture. This is categorized into two processes, slow (or s-) and rapid (or r-) process. In the former, neutron densities are typically low enough that the product of each neutron capture beta decays to a more stable isotope before capturing another neucleon. In r-process, however, the rate of neutron capture exceeds beta decay, alowing for the production of different, typically heavier, nuclei. As shown in the diagram, ligher elements are produced via the s-process and heavier via the r-process, with notable overlap in certain elements, such as Ba.\footnote{For the sake of this discussion, the light neutron-capture primary process elemnts (sometimes referred to as ``lighter element primary process") should be considered s-process elements. This different nucleosynthetic channel, first identified in \cite{Travaglio2004}, arose from an apparent disagreement between models of s-process enrichment in galactic chemical evolution studies andnd abundances of elements including Sr, Y, and Zr. However there is still uncertainty as to whether or not these differences can be explained with different models of weak s-process (which can occur in the winds of massive stars) \citep[e.g.][]{Prantzos2018}.} The s-process occurs predominately in the low-energy winds of low mass (1-8 M$_{\odot}$) AGB stars on time scales of 100's of Myr up to 1 Gyr. Exactly which elements are produced in a given AGB star varies with stellar mass and depends strongly on metallicity, which determines the available heavy seed nuclei. Sr, Y, Zr, and Ba, as well as C, N and F, are commonly used as tracers of this nucleosynthetic channel.

The dominant origin of r-process enrichment is highly uncertain, though possible channels include neutron star - neutron star and neutron star - black hole mergers, netrino driven winds in core collapse SNe, and exotic SNe, such as magnetic, jet-driven SNe, collapsars, hypernovae, and long-duration gamma-ray bursts \citep[see ][ for recent reviews]{Frebel2018, Cowan2019}. Eu is the most commonly used (and easy to observe) unambigious observational tracer of r-process enrichment; Ba is also often studied as a tracer, but requires care due to the significant contribution from s-process enricment. More detailed spectra \citep[e.g.][]{Ji2018a} can trace many more of these elements, which can be used as an important discriminator between nucleosynthetic models.

Finally, although the production of elements unaffected by the r-process are comparatively well understood, there are still large uncertainties in both nuclear reaction rates and stellar evolution properties such that there is generally significant variations in yields for any given element between groups who compute nucleosynthetic yield tables.
